### 2015-05-28 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:05:28	11:31:07	(Thu28.May)

Most modern active sonar imaging systems are usually far better at resolving details in range than in the angular direction. This is because it is relatively cheap and easy to collect a long and well sampled temporal representation of the incoming wave field, but hard to get the spatial equivalent. It would require a very large array with a lot of sensors, which is costly, complicated and impractical. An alternative is to synthesize a large array from a small array through a coherent combination of data from successive pings, a technique known as synthetic aperture sonar (SAS). Unfortunately, the difficult nature of navigating with high precision under water limits this technique to systems with advanced navigational equipment.

Due to the limitations imposed on hardware, a lot of research is put into improving the part of the digital image formation process. A core component here is the beamformer, a spatial band-pass filter used on the wave field to estimate each pixel value. A way to do this while allowing wide-band signals is to steer coarsely by applying delays to the sensor channels, followed by an individual weighting of them. Using complex weights, this implies fine adjustment of both the amplitude and the phase of the spatial filter. While conventional beamformers apply static weight functions, adaptive methods dynamically tailor the weights to the wave field.

@@

This article is outlined as follows: In Section \ref{methods}, we introduce the concept of adaptive beamforming and provide details on the MVDR and LCA method. Then, in \ref{maptogpu}, we investigate the complexity issue, discuss means for reducing arithmetic complexity, and detail an implementation that makes efficient use of the GPU's parallel resources. The final design is assessed in Sections \ref{images_and_benchmarks} and \ref{discussion}, where we provide benchmarks, comparisons with similar CPU implementations, and measures of how efficiently our implementation makes use of the GPU's resources.

@@

In modern active sonar imaging systems it is generally harder to obtain sufficient is to transmit and receive wideband acustic pulses using multi-sensor arrays. and process their record the backscattered wavefield using an array of hydrophones.  Through a coherent combination of the sensor data these

>>>>

One way to improve the performance of imaging devices is to make them more intelligent. 
To improve the performance of modern imaging devices is becoming increasingly intelligent, they adapt their behaviour to best fit the situation they are in. A regular compact camera for instance, adjusts its shutter opening and speed to

The image quality of any imaging device is mostly its by its hardware components, and . 
\IEEEPARstart{D}{ata-driven} methods have been introduced in various imaging fields over the years in attempts to improve image quality. One such method is the minimum variance distortionless response (MVDR) beamformer. Compared to conventional methods, MVDR is often able to improve image contrast and resolution, as shown in~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}.

Despite its inherent potential, the MVDR beamformer has yet to see widespread adoption in the active sonar field. There may be several reasons for this. For one, the method is not inherently robust, and may suffer from a phenomenon called signal cancellation in active systems~\cite{Widrow1982}. Another reason is that in its original form, the computational complexity is cubic with the number of channels, O($M^3$), while conventional beamformers are at O($M$). This is because a spatial covariance matrix is estimated and inverted for each image pixel.

@@

This article is outlined as follows:

% This article is outlined as follows: In Section \ref{methods}, we introduce the concept of adaptive beamforming and provide details on the MVDR and LCA method. Then, in \ref{maptogpu}, we investigate the complexity issue, discuss means for reducing arithmetic complexity, and detail an implementation that makes efficient use of the GPU's parallel resources. The final design is assessed in Sections \ref{images_and_benchmarks} and \ref{discussion}, where we provide benchmarks, comparisons with similar CPU implementations, and measures of how efficiently our implementation makes use of the GPU's resources.

@@

### 2015-05-28 ThesisJo/background.tex
la2pdf commit - 2015:05:28	11:31:07	(Thu28.May)

Consider a monochromatic plane wave impinging on an $M$-element array (\Fig{geometry}). Let the position of sensor $m$ be $\p_m$, th  of sensors at locations 
In this thesis we will relate all geometrical interpretations to the euclidean, spherical coordinate system shown in \Fig{geometry}. When assessing the imaging performance of a sonar its origin will be in the center of the array, but when assessing scenes its origin will be in the center of the image. A unit length vector $\vec u$ can be written as:
%
\begin{align*}
\vec u =
\begin{bmatrix}
\sin\theta \cos\phi \\
\sin\theta \sin\phi \\
\cos\theta
\end{bmatrix}
\end{align*}
%

Signal model. Monochromatic, plane wave:
%
\begin{align}
\s[n] = \begin{bmatrix}
         s_0[n] \\ s_1[n] \\ \vdots \\ s_{M-1}[n]
        \end{bmatrix}
      = \begin{bmatrix}
         Ae^{j(\Omega n - \k\T\p_0)} \\
         Ae^{j(\Omega n - \k\T\p_1)} \\
         \vdots \\
         Ae^{j(\Omega n - \k\T\p_{M-1})}
        \end{bmatrix}
      = Ae^{j\Omega n}\a(\k)
\end{align}
%
where we call $\a(\k)$ the signal's \emph{steering vector}.

>>>>

Consider a monochromatic plane wave impinging on an $M$-element array (\Fig{geometry}). Let the position of sensor $m$ be $\p_m$, th  of sensors at locations 
In this thesis we will relate all geometrical interpretations to the euclidean, spherical coordinate system shown in \Fig{geometry}. When assessing the imaging performance of a sonar its origin will be in the center of the array, but when assessing scenes its origin will be in the center of the image.

Signal model. Monochromatic, plane wave:
%
\begin{align}
\s[n] = \begin{bmatrix}
         s_0[n] \\ s_1[n] \\ \vdots \\ s_{M-1}[n]
        \end{bmatrix}
      = \begin{bmatrix}
         Ae^{j(\Omega n - \k\T\p_0)} \\
         Ae^{j(\Omega n - \k\T\p_1)} \\
         \vdots \\
         Ae^{j(\Omega n - \k\T\p_{M-1})}
        \end{bmatrix}
      = Ae^{j\Omega n}\a(\k)
\end{align}
%
where $A$ is the signal amplitude and $\Omega$ is normalized angular frequency. $\a(\k)$ is what we call signal's \emph{steering vector}.

A unit length vector $\vec u$ can be written as:
%
\begin{align*}
\vec u =
\begin{bmatrix}
\sin\theta \cos\phi \\
\sin\theta \sin\phi \\
\cos\theta
\end{bmatrix}
\end{align*}
%

\section{Active Sonar}

Most modern active sonar imaging systems are usually far better at resolving details in range than in the angular direction. This is because it is relatively cheap and easy to collect a long and well sampled temporal representation of the incoming wave field, but hard to get the spatial equivalent. It would require a very large array with a lot of sensors, which is costly, complicated and impractical. An alternative is to synthesize a large array from a small array through a coherent combination of data from successive pings, a technique known as synthetic aperture sonar (SAS). Unfortunately, the difficult nature of navigating with high precision under water limits this technique to systems with advanced navigational equipment.

Due to the limitations imposed on hardware, a lot of research is put into improving the part of the digital image formation process. A core component here is the beamformer, a spatial band-pass filter used on the wave field to estimate each pixel value. A way to do this while allowing wide-band signals is to steer coarsely by applying delays to the sensor channels, followed by an individual weighting of them. Using complex weights, this implies fine adjustment of both the amplitude and the phase of the spatial filter. While conventional beamformers apply static weight functions, adaptive methods dynamically tailor the weights to the wave field.
